---
title: "Aim 3 Dissertation R Analysis, Campylobacter Cattle and Human 
Comparative Analysis"
---

Barplot of ST and CC grouped by Cattle-derived and Human-derived isolates

```{r ST and CC Figures, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}

##Loading in libraries / installing packages
library(readxl)
library(dplyr)
library(tidyr)
library(tidyverse)
library(ggplot2)

###setting working directory to project data folder
#getwd()
#setwd(choose.dir())
#import data from RStudio fx

## These data were output from mlst, which queries our genomes with the PubMLST database ##
##setting new data variables to manpulate 

my_colors <- c("cattle_percent" = "#79bd8f", "human_percent" = "#70dfec")
dodge_width <- 0.7

human_st_cc <- human_datatable
cattle_st_cc <- cattle_datatable

human_st_sum <- human_st_cc %>%
  group_by(ST) %>%
  summarize(human_count = n()) %>%
  ungroup()

cattle_st_sum <- cattle_st_cc %>%
  group_by(ST) %>%
  summarize(cattle_count = n()) %>%
  ungroup()

st_data <- full_join(human_st_sum, cattle_st_sum, by = "ST")
st_data[is.na(st_data)] <- 0

st_data <- st_data %>%
  filter(cattle_count > 0)

total_human <- sum(st_data$human_count)
total_cattle <- sum(st_data$cattle_count)

st_data <- st_data %>%
  mutate(Rare = ifelse((human_count + cattle_count) == 1, "Rare ST", "NonRare"))

rare_sts <- st_data$ST[st_data$Rare == "Rare"]

print(rare_sts)

human_st_cc <- human_st_cc %>%
  mutate(ST = ifelse(ST %in% rare_sts, "Rare ST", as.character(ST)))

cattle_st_cc <- cattle_st_cc %>%
  mutate(ST = ifelse(ST %in% rare_sts, "Rare ST", as.character(ST)))

# Recreating the counts with Rare STs now as a countable group
human_st_sum2 <- human_st_cc %>%
  group_by(ST) %>%
  summarize(human_count = n()) %>%
  ungroup()

cattle_st_sum2 <- cattle_st_cc %>%
  group_by(ST) %>%
  summarize(cattle_count = n()) %>%
  ungroup()

st_data2 <- full_join(human_st_sum2, cattle_st_sum2, by = "ST")
st_data2[is.na(st_data2)] <- 0

# Calculate new percentages
st_data2 <- st_data2 %>%
  rowwise() %>%
  mutate(total = human_count + cattle_count,
         human_percent = ifelse(total == 0, 0, (human_count / total) * 100),
         cattle_percent = ifelse(total == 0, 0, (cattle_count / total) * 100)) %>%
  ungroup()
  
st_data2 <- st_data2 %>%
  filter(cattle_count > 0)

st_data_long <- st_data2 %>%
    pivot_longer(cols = c(human_percent, cattle_percent), names_to = 'Source', values_to = 'Percent')
  
st_data_long <- st_data_long %>%
    arrange(desc(total), .by_group = TRUE) %>%
    mutate(ST = factor(ST, levels = unique(ST)))
  
write.csv(st_data_long, "st_data_long_check.csv", row.names = FALSE)

ggplot(st_data_long, aes(x = ST, y = Percent, fill = Source)) +
  geom_bar(stat = 'identity', position = position_dodge(width = dodge_width)) +
  geom_text(aes(label = ifelse(Source == "human_percent", human_count, cattle_count)),
            position = position_dodge(width = dodge_width),
            vjust = -0.25,
            size = 3.5,
            color = "black") +
  scale_fill_manual(
    name = "Source",
    values = my_colors,
    labels = c("human_percent" = "Human Derived", "cattle_percent" = "Cattle Derived")
  ) +
  scale_y_continuous(limits = c(0,110), expand = c(0, 0)) +
  theme_minimal() +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    axis.line = element_line(color = "black")
  ) +
  labs(x = 'Sequence Type (ST)', y = 'Percentage') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


##Counts and building plots for CC data ##

human_cc_sum <- human_st_cc %>%
  group_by(CC) %>%
  summarize(human_count = n()) %>%
  ungroup()

cattle_cc_sum <- cattle_st_cc %>%
  group_by(CC) %>%
  summarize(cattle_count = n()) %>%
  ungroup()

cc_data <- full_join(human_cc_sum, cattle_cc_sum, by = "CC")

cc_data <- cc_data %>%
  rowwise() %>%
  mutate(total = human_count + cattle_count, 
         human_percent = ifelse(total == 0, 0, (human_count/ total)*100),
         cattle_percent = ifelse(total == 0, 0, (cattle_count/ total) *100)) %>%
  ungroup()

cc_data <- cc_data %>%
  arrange(desc(total))


cc_data[is.na(cc_data)] <- 0

cc_data_long <- cc_data %>%
  pivot_longer(cols = c(human_percent, cattle_percent), names_to = 'Source', values_to = 'Percent')

cc_data_long[is.na(cc_data_long)] <- 0

cc_data_long$CC <- as.character(cc_data_long$CC)

cc_data_long$CC <- case_when(
  cc_data_long$CC == "0" ~ "CC unassigned",
  TRUE ~ cc_data_long$CC
)

cc_data_long <- cc_data_long %>%
  filter(cattle_count > 0 )

#Plotting!#
ggplot(cc_data_long, aes(x = CC, y = Percent, fill = Source)) +
  geom_bar(stat = 'identity', position =  position_dodge(width = dodge_width)) +
  geom_text(aes(y = Percent, label = ifelse(Source == "human_percent", human_count, cattle_count)),
            position = position_dodge(width = dodge_width),
            vjust = -0.25,
            size = 3.5, 
            color = "black") +
  scale_fill_manual(
    name = "Source",
    values = my_colors,
    labels = c("human_percent" = "Human Derived", "cattle_percent" = "Cattle Derived")
  ) +
  scale_y_continuous(limits = c(0,110), expand = c(0,0)) +
  theme_minimal() +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    axis.line = element_line(color = "black")
  ) + 
  labs(x = 'Clonal Complex (CC)', y = 'Percentage') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
Venndiagram of cattle vs. human derived genes

```{r Cattle vs Human derived genes, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}

library(readxl)
library(dplyr)
library(tidyr)
library(tidyverse)
library(ggplot2)
library(ggvenn)
library(stringr)

###setting working directory to project data folder
#getwd()
#setwd(choose.dir())
#import data from RStudio fx

## These data were output from Panaroo ##

my_colors <- c(Cattle = "#79bd8f", Human = "#70dfec")

gene_data <- gene_presence_absence
human_isolates <- human_datatable
cattle_isolates <- cattle_datatable

human_genes <- gene_data %>%
  select(c('Gene', human_isolates$Isolates)) %>%
  filter(rowSums(!is.na(.)) > 1) %>%
  pull(Gene)

cattle_genes <- gene_data %>%
  select(c('Gene', cattle_isolates$Isolates)) %>%
  filter(rowSums(!is.na(.)) > 1) %>%
  pull(Gene)

list_genes <- list(Human = human_genes, Cattle = cattle_genes)
ggvenn_plot <- ggvenn(list_genes)
##Plotting

ggvenn_plot + 
  scale_fill_manual(values = my_colors) +
  theme_minimal() +
  labs(title = "Venn Diagram of Gene Presence in Human and Cattle Isolates")
#Rough Venndiagram was then edited in Inkscape


##Generating table of genes that are population specific (cattle derived vs. human derived isolates) 

human_isolatesID <-human_isolates$Isolates
cattle_isolatesID <- cattle_isolates$Isolates

long_gene_data <- gene_data %>%
  pivot_longer(cols = -c(Gene, Annotation), names_to = "Isolates", values_to = "Presence", values_drop_na = TRUE)

human_specific_genes <- long_gene_data %>%
  filter(Isolates %in% human_isolatesID) %>%
  anti_join(long_gene_data %>% filter(Isolates %in% cattle_isolatesID), by = "Gene")

cattle_specific_genes <- long_gene_data %>%
  filter(Isolates %in% cattle_isolatesID) %>%
  anti_join(long_gene_data %>% filter(Isolates %in% human_isolatesID), by = "Gene")

human_specific_genes <- human_specific_genes %>%
  left_join(human_isolates, by = "Isolates")

cattle_specific_genes <- cattle_specific_genes %>%
  left_join(cattle_isolates, by = "Isolates")

combined_genes <- bind_rows(human_specific_genes, cattle_specific_genes) %>%
  group_by(Gene, Annotation) %>%
  summarise(
    Population = ifelse(first(Isolates) %in% human_isolatesID, "Human Derived", "Cattle Derived"),
    Isolates = toString(unique(Isolates)),
    STs = toString(unique(ST)),
    CCs = toString(unique(na.omit(CC))),
    Frequency = n(),
    .groups = 'drop'
  )

write.csv(combined_genes, "unique_population_genes.csv", row.names = FALSE)

```
Looking at the number of accessory genes per isolates by cattle and human derived isolates. 
Core genes were defined as in 99% or > of isolates and 1320 genes were present in total

```{r Violin Plot & Statistic, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}


##Loading in libraries / installing packages
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggbeeswarm)

###setting working directory to project data folder
#getwd()
#setwd(choose.dir())
#import data from RStudio fx

## These data were generated from Panaroo output ###

my_colors <- c(Cattle = "#79bd8f", Human = "#70dfec")

gene_data <- gene_presence_absence
human_isolates <- human_datatable$Isolates
cattle_isolates <- cattle_datatable$Isolates

gene_data <- gene_data %>%
  mutate(Total_Isolates = rowSums(select(., -Gene))) %>%
  mutate(Percent_Presence = Total_Isolates / (ncol(.) - 2))

gene_data <- gene_data %>%
  mutate(Category = case_when(
    Percent_Presence >= 0.99 ~ "Core", 
    TRUE ~ "Accessory"
  ))

core_gene_count <- sum(gene_data$Category == "Core")
cat("Number of core genes:", core_gene_count, "\n")
#Yay! 1320 core genes were present same as Panaroo output!

accessory_genes <- gene_data %>%
  filter(Category == "Accessory") %>%
  pivot_longer(cols = -c(Gene, Total_Isolates, Percent_Presence, Category), names_to = "Isolates", values_to = "Presence")


accessory_gene_counts <- accessory_genes %>%
  mutate(Source = case_when(
    Isolates %in% human_isolates ~ "Human",
    Isolates %in% cattle_isolates ~ "Cattle"
  )) %>%
  group_by(Source, Isolates) %>%
  summarise(Accessory_Gene_Count = sum(Presence), .groups = 'drop')
  
            
ggplot(accessory_gene_counts, aes(x = Source, y = Accessory_Gene_Count, fill = Source)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.2, outlier.shape = NA) +
  geom_point(position = position_jitter(width = 0.2), shape = 21, size = 3, color = "black", fill = "white") +
  scale_fill_manual(values = my_colors) +
  labs(title = "Accessory Gene Counts in Human vs. Cattle Isolates",
       x = "Source",
       y = "Accessory Gene Count") +
  theme_minimal() +
  ylim(0, 700) 

summary_stats <- accessory_gene_counts %>%
  group_by(Source) %>%
  summarise(
    Median = median(Accessory_Gene_Count),
    First_Quartile = quantile(Accessory_Gene_Count, 0.25),
    Third_Quartile = quantile(Accessory_Gene_Count, 0.75)
  )

print(summary_stats)

stats <- wilcox.test(Accessory_Gene_Count ~ Source, data = accessory_gene_counts)
print(stats)


t_test_result <- t.test(Accessory_Gene_Count ~ Source, data = accessory_gene_counts)
t_test_result


contingency_table <- accessory_genes %>%
  mutate(Source = case_when(
    Isolates %in% human_isolates ~ "Human",
    Isolates %in% cattle_isolates ~ "Cattle"
  )) %>%
  group_by(Gene, Source) %>%
  summarise(Presence = sum(Presence), .groups = 'drop') %>%
  pivot_wider(names_from = Source, values_from = Presence, values_fill = list(Presence = 0))

# Perform chi-square test
chi_square_result <- chisq.test(contingency_table[, -1])
print(chi_square_result)



```
AMR determinants by group cattle vs. human

```{r AMR, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}

library(tidyverse)
library(dplyr)
library(ggplot2)
library(forcats)

###setting working directory to project data folder
#getwd()
#setwd(choose.dir())
#import data from RStudio fx
### These data were output from quering genomes with AMRFINDER PLUS ###

amr_data <- amrfinderplus_campy
cattle_isolates <- cattle_datatable$Isolates
human_isolates <- human_datatable$Isolates

my_colors <- c(Cattle = "#79bd8f", Human = "#70dfec")

#Density plot for the # of resistance determinants 
amr_summary <- amr_data %>%
  mutate(Source = case_when(
    AccessionNum %in% human_isolates ~ "Human",
    AccessionNum %in% cattle_isolates ~ "Cattle"
  )) %>%
  group_by(AccessionNum, Source) %>%
  summarise(Resistance_Count = n(), .groups = 'drop')

write.csv(amr_summary, "amr_summary.csv", row.names = FALSE)

resistance_count <- ggplot(amr_summary, aes(x = Resistance_Count,  fill = Source)) +
  geom_density(alpha = 0.5) +
  scale_x_continuous(limits = c(0,10), breaks =1:10) +
  scale_y_continuous(limits= c(0.0, 0.7))+
  scale_fill_manual(values = my_colors) +
  labs(title = "Density of Antibiotic Resistance Determinants",
       x = "Number of Antibiotic Resistance Determinants",
       y = "Density") +
  theme_minimal()

print(resistance_count)

ggsave("arg_count.svg", plot = resistance_count, device = "svg")

#Ceating bar plots of the mechanisms of resistance
resistance_summary <- amr_data %>%
  mutate(Source = case_when(
    AccessionNum %in% human_isolates ~ "Human",
    AccessionNum %in% cattle_isolates ~ "Cattle"
  )) %>%
  group_by(AccessionNum, Source) %>%
  group_by(Class, Gene_symbol, Source) %>%
  summarise(Count = n(), .groups = 'drop')

resistance_summary <- resistance_summary %>%
  mutate(Gene_symbol = fct_reorder(Gene_symbol, Count, .fun = sum))

write.csv(resistance_summary, "resistance_summary.csv", row.names = FALSE)
#creating a seperate graph for beta-lactams as there are a variety of mechanisms
ggplot(subset(resistance_summary, Class == "BETA-LACTAM"), #Replace here for other graphs 
       aes(x = Count, y = Gene_symbol, fill = Source)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = my_colors) +
  labs(title = "Specific Antibiotic Resistance by Mechanism",
       x = "Number of Isolates",
       y = "Mechanism") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, hjust = 1))

#One graph for tetracycline, quinolone and macrolides here
selected_classes <- c("TETRACYCLINE", "QUINOLONE", "MACROLIDE")

resistance_summary_x2 <- resistance_summary %>%
  filter(Class %in% selected_classes) %>%
  mutate(Class_Mechanism = paste(Class, Gene_symbol, sep = ": ")) %>%
  mutate(Class_Mechanism = fct_reorder(Class_Mechanism, Count, .fun = sum))
         
ggplot(resistance_summary_x2, aes(y = Class_Mechanism, x = Count, fill = Source)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = my_colors) +
  scale_x_continuous(limits = c(0,150)) +
  labs(title = "Antibiotic resistance by Mechanism and Class",
        x = "Number of Isolates",
        y =  "Mechanism")+
  theme_minimal() + 
  theme(axis.text.y = element_text(angle = 0, hjust = 1))

#Got a table of percent based on the total # of amr mechanisms identified

resistance_summary_x3 <- amr_data %>%
  mutate(Source = case_when(
    AccessionNum %in% human_isolates ~ "Human",
    AccessionNum %in% cattle_isolates ~ "Cattle"
  )) %>%
  group_by(Class, Gene_symbol, Source) %>%
  summarise(Count = n(), .groups = 'drop')

total_counts <- resistance_summary_x3 %>%
  group_by(Source) %>%
  summarise(Total = sum(Count))

resistance_summary_x3 <- resistance_summary_x3 %>%
  left_join(total_counts, by = "Source") %>%
  mutate(Percentage = Count/ Total * 100)

write.csv(resistance_summary_x3, "percent_positive_amr_mechanism.csv", row.names = FALSE)

#here graphing % of isolates that are positive for a particular resistance determinant within a class and grouped by Source
#i.e. cattle vs. human derived isolates

resistance_summary_class <- amr_data %>%
  mutate(Source = case_when(
    AccessionNum %in% human_isolates ~ "Human",
    AccessionNum %in% cattle_isolates ~ "Cattle"
  )) %>%
  group_by(Class, Source) %>%
  summarise(Count = n(), .groups = 'drop')

total_counts_class <- resistance_summary_class %>%
  group_by(Class) %>%
  summarise(Total = sum(Count), .groups = 'drop')

resistance_summary_class <- resistance_summary_class %>%
  left_join(total_counts_class, by = "Class") %>%
  mutate(Percentage = Count/ Total * 100)

write.csv(resistance_summary_class, "percent_positive_per_class.csv", row.names = FALSE)

ggplot(resistance_summary_class, aes(x = Class, y = Percentage, fill = Source)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = my_colors) +
  labs(
    title = "Percentage of Each  Resistance Class by Source",
    x = "Class",
    y = "Percentage (%)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

amr_summary_classes <- amr_data %>%
  mutate(Source = case_when(
    AccessionNum %in% human_isolates ~ "Human",
    AccessionNum %in% cattle_isolates ~ "Cattle"
  )) %>%
  group_by(AccessionNum, Source) %>%
  summarise(Resistance_classes = n_distinct(Class), .groups = 'drop')

write.csv(amr_summary_classes, "amr_summary_classses.csv", row.names = FALSE)

mdr_isolates <- amr_summary_classes %>%
  filter(Resistance_classes > 3)

print(mdr_isolates)

amr_data_mdr <- amr_data %>%
  filter(AccessionNum %in% mdr_isolates$AccessionNum)

write.csv(amr_data_mdr, "amr_data_mdr.csv", row.names = FALSE)

arg_resistance_classes <- ggplot(amr_summary_classes, aes(x = Resistance_classes, fill = Source)) +
  geom_density(alpha = 0.5) +
  scale_x_continuous(limits = c(0, 10), breaks = 0:10) +
  scale_fill_manual(values = my_colors) +
  labs(
    title = "Density of Antibiotic Resistance Classes",
    x = "Number of Antibiotic Resistance Classes",
    y = "Density"
  ) +
  theme_minimal()

ggsave("arg_classes.svg", plot = arg_resistance_classes, device = "svg")
```
Silhouette analysis of the hierarachial clustering for National/Global analysis:
  
These data were generated from BIGSdb through PubMLST functionality. ST-982  human and cattle genomes (from Michigan - our studies) were 
uploaded to PubMLST a query was run based on the sequence type. cgMLST profiles and minimum spanning tree was generated
based on the cgMLST profiles. ReproTree was used to generate clusters via hierachial clustering on the cgMLST hamming distance derived from cgmlst-dist
which is implemented in ReporTree. Here we are using the clustering assignment and the hamming distance matrix from cg-mlst dists to conduct a Silhouette 
analysis. Clusters with the highest Silouette (max) and negative Silouette area (min) will be selected. Additional data from ReproTree will be used to identify areas 
of cluster stability. The cluster selected based on cluster stability, and silouette analysis  will then be visualized with the cgMLST minimum spanning tree. 
These defined clusters will then be used in futher analysis to evaluate phylogeography and epidemiologic/statisicial associations based on the cgMLST HC within ST-982 .

```{r Clusters, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
library(tidyverse)
library (dplyr)
library(cluster)
###setting working directory to project data folder
#getwd()
#setwd(choose.dir())
#import data from RStudio fx

#for the distance matrix need the row names to be the first column (not 1, 2, 3 etc.)
dist_data <- BIGSdb_dist_hamming
colnames(dist_data) <- sub("^X", "", colnames(dist_data))

sample_ids <- rownames(dist_data)
write.csv(sample_ids, "sample_ids.csv", row.names = FALSE)

#dist_matrix cleaves rows and column names
dist_matrix <- as.dist(dist_data)

#cleaning and reordering clustering data to match cleaned distance matrix
cluster_data <- BIGSdb_clusterComposition
colnames(cluster_data) <- sub("^X.", "", colnames(cluster_data))

#ensuring unique cluster_ids in cluster id including singletons
all_partitions_transformed <- data.frame()

cluster_id_counter <- 0

for(partition in unique(cluster_data$partition)) {
  partition_data <- cluster_data[cluster_data$partition == partition, ]
  
  partition_data <- partition_data %>%
    mutate(ordering_id = as.numeric(sub(".*_([0-9]+)$", "\\1", cluster)))
  
  partition_data <- partition_data %>%
    arrange(ordering_id)
  
  partition_data <- partition_data %>%
    mutate(cluster_id = match(cluster, unique(cluster)))
  
  partition_long <- partition_data %>%
    separate_rows(samples, sep = ",") %>%
    mutate(samples = trimws(samples))

  all_partitions_transformed <- rbind(all_partitions_transformed, partition_long)
}

write.csv(all_partitions_transformed, "all_partitions_transformed.csv", row.names = FALSE)

#slightly computationally intensive step of ordering 
#here reordering the clustering data to match the data in the distance matrix (sample_ids variable)
  
unique_partitions <- unique(all_partitions_transformed$partition)
total_partitions <- length(unique_partitions)

output_file_path <- "all_partitions_reordered.csv"  

for(partition in unique(all_partitions_transformed$partition)) {
  partition_data <- all_partitions_transformed %>%
    filter(partition == partition, samples %in% sample_ids) %>%
    arrange(match(samples, sample_ids))
  
  if(partition == unique(all_partitions_transformed$partition)[1]) {
    write.csv(partition_data, file = output_file_path, row.names = FALSE)
  } else {
    write.csv(partition_data, file = output_file_path, row.names = FALSE, append = TRUE)
  }
}

#re-import the reordered csv here...

cluster_data_final <- all_partitions_reordered

##Here we are running the silhouette analysis on each of the clusters identified in the data
winning_interation <- NULL
max_silhouette_avg <- -Inf
min_neg_silhouette_area <- -Inf

silhouette_results <- data.frame(
    Partition = character(),
    AvgSilhouetteScore = numeric(),
    NegSilhouetteArea = numeric(),
    stringsAsFactors = FALSE
  )
  
  for (i in seq_along(unique_partitions)) {
    partition <- unique_partitions[i]
    
    # Report progress
    cat(sprintf("Analyzing partition %s (%d of %d)\n", partition, i, total_partitions))
    
    current_partition_data <- cluster_data_final[cluster_data_final$partition == partition, ]
 
    # Check if there are at least 2 different clusters
    if (length(unique(current_partition_data$cluster_id)) < 2) {
      cat("Not enough clusters in partition", partition, "to calculate silhouette scores.\n")
      next  # Skip this iteration
    }
    # Perform silhouette analysis
    silhouette_scores <- silhouette(current_partition_data$cluster_id, dist_matrix)
    avg_silhouette_score <- mean(silhouette_scores[, "sil_width"], na.rm = TRUE)
    neg_silhouette_area <- sum(silhouette_scores[silhouette_scores[, "sil_width"] < 0, "sil_width"], na.rm = TRUE)
    
    if (avg_silhouette_score > max_silhouette_avg || neg_silhouette_area < min_neg_silhouette_area) {
      winning_interation <- partition
      max_silhouette_avg <- avg_silhouette_score
      min_neg_silhouette_area <- neg_silhouette_area
    }
    silhouette_results <- rbind(silhouette_results, data.frame(
      Partition = partition,
      AvgSilhouetteScore = avg_silhouette_score,
      NegSilhouetteArea = neg_silhouette_area
    ))
  }

write.csv(silhouette_results, "silhouette_analysis_results_outliers_excluded.csv", row.names = FALSE)
##based on the stability regions and clustering data here, partition or iteration #94 was chosen
## criteria included partition where >1 cluster had >1 isolate, and in a stability region AND Silhouette analysis
````
ReproTree, ST-982 clustering by HHS regions

```{r Clusters vs. region analysis, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}

library(tidyverse)
library (dplyr)
library(broom)
library(purrr)
library(tidyr)
library(ggplot2)
###setting working directory to project data folder
#getwd()
#setwd(choose.dir())
#import data from RStudio fx


#Defining US State Region (HHS Regions)
region_1 <- c("CT", "ME", "MA", "NH", "RI", "VT" )
region_2 <- c("NJ", "NY","PR", "VI")
region_3 <- c("DE", "DC", "MD", "PA","VA", "WV")
region_4 <- c("AL", "FL", "GA", "KY", "MS", "NC", "SC", "TN")
region_5 <- c("IL", "IN", "MI", "MN", "OH", "WI")
region_6 <- c("AR", "LA", "NM", "OK", "TX")
region_7 <- c("IA", "KS", "MO", "NE")
region_8 <- c("CO", "MT", "ND", "SD", "UT", "WY")
region_9 <- c("AZ", "CA", "HI", "NV", "AS", "MP", "TT")
region_10 <- c("AL", "ID", "OR", "WA")
#HHS regions: https://www.hhs.gov/about/agencies/iea/regional-offices/index.html

my_colors <- c('Cluster 1'= '#63BB70', 'Cluster 2' = '#7CCEFD', 'Cluster 3' = '#FF5252', 'Cluster 4' = '#9574CD', 'Singleton' = '#58585A', 'Cluster 5' = '#FFF175', 'Cluster 6' = '#F44ED8', 'Cluster 7' = '#F77D55')

#selected single-94x1.0 partition from previous cluster analysis (above)
# Only using a sub-set of complete data here and in subsequent analyses below... n=723 
cluster_data_final <- all_partitions_reordered
selected_parition <- cluster_data_final %>%
  filter(partition == "single-94x1.0")

metadata <- metadata_global

combined_data <- metadata %>%
                 left_join(selected_parition %>% select(samples, cluster_id), by = c("ID" = "samples"))

#cleaning metadata from isolates
combined_data <- combined_data %>%
  mutate(region = ifelse(grepl("Michigan", region), "MI", region))

combined_data <- combined_data %>%
  mutate(region = ifelse(grepl("Georgia", region), "GA", region))

combined_data <- combined_data %>%
  mutate (US_region = case_when(
    region %in% region_1 ~ "Region 1",
    region %in% region_2 ~ "Region 2",
    region %in% region_3 ~ "Region 3",
    region %in% region_4 ~ "Region 4",
    region %in% region_5 ~ "Region 5",
    region %in% region_6 ~ "Region 6",
    region %in% region_7 ~ "Region 7",
    region %in% region_8 ~ "Region 8",
    region %in% region_9 ~ "Region 9",
    region %in% region_10 ~ "Region 10",
  ))  

#Removing NA georgraphic location samples
clean_combined_data <- combined_data %>%
  filter(!is.na(cluster_id) & !is.na(US_region))

clean_combined_data <- clean_combined_data %>%
  add_count(cluster_id) %>%
  mutate(cluster_id = ifelse(n == 1, "Singleton", as.character(cluster_id))) %>%
  select(-n)  

#Renaming clusters cluster 1 -> 7 + Singletons

cluster_ranks <- clean_combined_data %>%
  filter(cluster_id != "Singleton") %>%
  count(cluster_id) %>%
  arrange(desc(n)) %>%
  mutate(rank = row_number(), 
         new_cluster_id = paste("Cluster", rank)) %>%
  select(-n)

clean_combined_data <- clean_combined_data %>%
  left_join(cluster_ranks, by = "cluster_id") %>%
  mutate(cluster_id = ifelse(cluster_id != "Singleton", new_cluster_id, cluster_id)) %>%
  select(-rank, -new_cluster_id)

write.csv(clean_combined_data, "clean_combined_data.csv", row.names = FALSE)

#summarizing data for charts, table and analysis
region_cluster_summary <- clean_combined_data %>%
  group_by(US_region, cluster_id) %>%
  summarise( count = n()) %>%
  ungroup()

region_totals <- clean_combined_data %>%
  group_by(US_region) %>%
  summarise(total_samples = n())

region_cluster_summary <- clean_combined_data %>%
  group_by(US_region, cluster_id) %>%
  summarise(cluster_samples = n()) %>%
  left_join(region_totals, by = "US_region") %>%
  mutate(percentage = (cluster_samples / total_samples) * 100) %>%
  ungroup()

region_cluster_summary <- region_cluster_summary %>%
  mutate(percentage_display = paste0(round(percentage, 2), "% (n = ", cluster_samples, ")"))

region_cluster_summary

write.csv(region_cluster_summary, "region_cluster_summary.csv", row.names = FALSE)

#function to create piecharts 
 create_pie_chart_for_region <- function(region_name, data) {
   region_data <- filter(region_cluster_summary, US_region == region_name)
   ggplot(region_data, aes(x = "", y = cluster_samples, fill = as.factor(cluster_id))) +
     geom_bar(width = 1, stat = "identity", color = "black") +
     geom_text(aes(label = percentage_display),
               color = "white", size = 3) +
     coord_polar("y", start = 0) +
     scale_fill_manual(values = my_colors) +
     theme_void() +
     labs(title = paste("Cluster Distribution in", region_name), fill = "Cluster ID") +
     guides(fill = guide_legend(title = "Cluster ID"))
 }

#change [name] here to get pie chart from the different regions
create_pie_chart_for_region( ["Region 1"], region_cluster_summary)


```
Generating results region vs. cluster

```{r odds ration for regions vs cluster, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}

library(ggplot2)
library(dplyr)
library(tidyverse)
library(forcats)

#[import clean_combined_data]

unique_clusters <- unique(clean_combined_data$cluster_id)
unique_regions <- unique(clean_combined_data$US_region)

combinations <- expand.grid(cluster = unique_clusters, region = unique_regions)

#Function to caclulate odds ratio##
calculate_odds_ratio <- function(cluster, region, data) {
 
  cluster_presence <- data$cluster_id == cluster
  region_presence <- data$US_region == region
  
  contingency_table <- table(cluster_presence, region_presence)
  
  if (all(dim(contingency_table) == c(2, 2))) {
    fisher_test <- fisher.test(contingency_table)
    odds_ratio <- fisher_test$estimate
    conf.low <- fisher_test$conf.int[1]
    conf.high <- fisher_test$conf.int[2]
    p.value <- fisher_test$p.value
  } else {
    # Assign NA if table is not 2x2
    odds_ratio <- NA
    conf.low <- NA
    conf.high <- NA
    p.value <- NA
  }
  
  # Return results
  return(data.frame(cluster = cluster, region = region, 
                    odds_ratio = odds_ratio, 
                    conf.low = conf.low, conf.high = conf.high, 
                    p.value = p.value))
}



results <- purrr::pmap_df(combinations, ~ calculate_odds_ratio(..1, ..2, clean_combined_data))

# Benjamini-Hochberg correction for p-values
results <- results %>%
  mutate(p.value.adj = p.adjust(p.value, method = "BH"))


results
write.csv(results, "results_region_vs_cluster.csv", row.names = FALSE)

#Created a table.
```
Evaluation of cluster vs. time

```{r Clusters v. time, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}

library(tidyverse)
library(dplyr)
library(ggplot2)

filtered_data_time <- combined_data %>%
  group_by(cluster_id) %>%
  filter(n() >= 10) %>%
  ungroup()

time_summary <- filtered_data_time %>%
  group_by(year, cluster_id) %>%
  summarise(count = n()) %>%
  ungroup()

ggplot(time_summary, aes(x = year, y = count, group = cluster_id, color = cluster_id)) +
  geom_line() +
  labs(x = "Year", y = "Count", title = "Time Trend of Each Cluster") +
  theme_minimal()+
  ylim(0,100)

binary_data <- filtered_data_time %>%
  mutate(presence = 1) %>%
  complete(year, cluster_id, fill = list(presence = 0))

# Logistic regression for each cluster
trend_results <- binary_data %>%
  group_by(cluster_id) %>%
  do(tidy(glm(presence ~ year, data = ., family = binomial()))) %>%
  ungroup()

# View trend results
trend_results
write.csv(trend_results, "trend_results_time.csv", row.names = FALSE)

###No changes overtime from each of the clusters

```
genome size vs. time/cluster ; evaluation of genome size. The hypothesis here is that ST-982 might represent 
another evolving lineage similar to CC-61 which is a cattle associated lineage that has show specialization through gene loss. 
ST-982 is closely related to CC-61 phylogenetically. Therefore specialization will be assessed by average genome size per year

```{r genomes size vs. time & time, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}

library(tidyverse)
library(ggplot2)
library(dplyr)


genome_data <- BIGSdb_genome_data
meta_data <- clean_combined_data

genome_data_1<- genome_data %>% select(Isolate, Total.length)
meta_data_1 <- meta_data %>% select(isolate, year)

meta_data_1 <- meta_data_1 %>% rename(Isolate = isolate)

merged_data <- merge(genome_data_1,meta_data_1, by="Isolate")

merged_data <- merged_data %>%
              filter(!is.na(Total.length) &!is.na(year))

filtered_data <- merged_data %>%
  group_by(year) %>%
  filter(n() > 50) %>%
  ungroup()

average_genome_size <- filtered_data %>% 
  group_by(year) %>% 
  summarize(average_size = mean(Total.length))

model <- lm(average_size ~ year, data = average_genome_size)
summary(model)

ggplot(average_genome_size, aes(x = year, y = average_size)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Average Genome Size Over Time",
       x = "Year",
       y = "Average Genome Size")
# No relationship between average_genome_size and year...
#genome size does not differ don't have data on # of coding sequences but 
# that could be a more sensitive test ...

genome_data_2<- genome_data %>% select(Isolate, Total.length)
meta_data_2 <- meta_data %>% select(isolate, cluster_id)

meta_data_2 <- meta_data_2 %>% rename(Isolate = isolate)

merged_data_2 <- merge(genome_data_2,meta_data_2, by="Isolate")

merged_data_2 <- merged_data_2 %>%
  filter(!is.na(Total.length) &!is.na(cluster_id))

filtered_data_2 <- merged_data_2 %>%
  group_by(cluster_id) %>%
  filter(n() > 50) %>%
  ungroup()

filtered_data_2 <- filtered_data_2 %>%
  filter(cluster_id %in% c("Cluster 1", "Cluster 2"))

t_test_result <- t.test(Total.length ~ cluster_id, data = filtered_data_2)
t_test_result

my_colors <- c('Cluster 1'= '#63BB70', 'Cluster 2' = '#7CCEFD')

ggplot(filtered_data_2, aes(x = cluster_id, y = Total.length, fill = cluster_id)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.2, outlier.shape = NA) +
  geom_point(position = position_jitter(width = 0.2), shape = 21, size = 3, color = "black", fill = "white") +
  scale_fill_manual(values = my_colors) + 
  labs(title = "Genome Size Distribution by Cluster",
       x = "Cluster ID",
       y = "Genome Size") +
  theme_minimal() +
  ylim(1400000, 2000000) 

summary_stats <- filtered_data_2 %>%
  group_by(cluster_id) %>%
  summarise(
    Median = median(Total.length),
    First_Quartile = quantile(Total.length, 0.25),
    Third_Quartile = quantile(Total.length, 0.75)
  )

print(summary_stats)

stats <- wilcox.test(Total.length ~ cluster_id, data = filtered_data_2)
print(stats) #Wilcox since the data is not normally distributed

##Cluster 2 has bigger genomes! interesting! No relationship between time and cluster or time and genome size...
# See above. Could point to diversification/acquisition of MGEs ?
```
Measures of alpha diversity v. time/region/cluster
Here looking at cg-mlst profiles and to determine diversity within Cluster 1 and 2 and cgmlst profile diversity
across time and region. Using 436 alleles to assess diversity that were present in the full dataset 1152 ST-982 isolates
then narrowed that to 723 isolates that have complete metadata. 

```{r alpha diversity v. time/region/cluster, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}

#clean combined (ID, year, cluster_id and US_region)
# Again only looking at "complete" records here...

library(vegan)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(tidyr)


#Alpha Diversity...
meta_data <-clean_combined_data
cgmlst <- BIGSdb_profiles
#import with "-" as NA or Missing...

na_columns<- cgmlst %>%
  select(-1) %>%
  summarise_all(~sum(is.na(.))) %>%
  select_if(~any(.)) %>%
  names()

#reimport with cgmlst rownames as column 1 for later merge to work...
cgmlst <- BIGSdb_profiles
#707 alleles have some missing data...

cgmlst_clean <- cgmlst %>%
  select(-all_of(na_columns))

#436 alleles assessed with complete data across the entire dataset...

shannon_diversity <- diversity(cgmlst_clean, index = "shannon")
simpson_diversity <- diversity(cgmlst_clean, index = "simpson")



alpha_diversity <- data.frame(Isolate_ID = rownames(cgmlst_clean),
                              shannon = shannon_diversity, simpson = simpson_diversity)


meta_data <- meta_data %>% select(ID, year, cluster_id, US_region)

meta_data <- meta_data %>% rename(Isolate_ID = ID)

merged_data <- merge(meta_data,alpha_diversity, by= "Isolate_ID")

merged_data <- merged_data %>%
  filter(!is.na(year) &!is.na(US_region))

write.csv(merged_data, "alpha_diversity.csv", row.names = FALSE)

#stats and visualization for alpha diversity metrics


## Alpha diversity vs. Time
merged_data$year <- as.factor(merged_data$year)

year_count <- merged_data %>% group_by(year) %>% summarise(count = n())
years_to_keep <- year_count %>% filter(count >= 10) %>% select(year)

alpha_diversity_final <- merged_data %>% filter(year %in% years_to_keep$year)

shannon_aov <- aov(shannon ~ year, data = alpha_diversity_final)
summary(shannon_aov)

TukeyHSD(shannon_aov) #<- post-hoc analysis for the ANOVA

simpson_aov <- aov(simpson ~ year, data = alpha_diversity_final)
summary(simpson_aov)

TukeyHSD(simpson_aov) 

ggplot(alpha_diversity_final, aes(x = year, y = shannon, group = year)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Shannon Diversity Index Over Time",
       x = "Year",
       y = "Shannon Index") +
  ylim(2,6)
  
  
ggplot(alpha_diversity_final, aes(x = year, y = simpson, group = year)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Simpson Diversity Index Over Time",
       x = "Year",
       y = "Simpson Index") +
  ylim(0.7,1)

#Alpha Diversity vs. Cluster
cluster_count <- alpha_diversity_final %>% group_by(cluster_id) %>% summarise(count = n())
clusters_to_keep <- cluster_count %>% filter(count >= 10) %>% select(cluster_id)

alpha_diversity_cluster <- alpha_diversity_final %>% filter(cluster_id %in% clusters_to_keep$cluster_id)

shannon_aov_cluster <- aov(shannon ~ cluster_id, data = alpha_diversity_cluster)
summary(shannon_aov_cluster)
TukeyHSD(shannon_aov_cluster)

simpson_aov_cluster <- aov(simpson ~ cluster_id, data = alpha_diversity_cluster)
summary(simpson_aov_cluster)
TukeyHSD(simpson_aov_cluster) 


ggplot(alpha_diversity_cluster, aes(x = cluster_id, y = shannon, group = cluster_id)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Shannon Diversity Index by Cluster",
       x = "Cluster ID",
       y = "Shannon Index") +
  ylim(2,6)

ggplot(alpha_diversity_cluster, aes(x = cluster_id, y = simpson, group = cluster_id)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Simpson Diversity Index by Cluster",
       x = "Cluster ID",
       y = "Simpson Index") +
  ylim(0.7,1)

#Alpha Diversity vs. Region
region_count <- alpha_diversity_final %>% group_by(US_region) %>% summarise(count = n())
regions_to_keep <- region_count %>% filter(count >= 10) %>% select(US_region)

alpha_diversity_region <- alpha_diversity_final %>% filter(US_region %in% regions_to_keep$US_region)

# ANOVA for Shannon Index by region
shannon_aov_region <- aov(shannon ~ US_region, data = alpha_diversity_region)
summary(shannon_aov_region)

# ANOVA for Simpson Index by region
simpson_aov_region <- aov(simpson ~ US_region, data = alpha_diversity_region,
                          
summary(simpson_aov_region)

region_order <- c("Region 1", "Region 2", "Region 3", "Region 4", "Region 5", "Region 6",
                  "Region 7", "Region 8", "Region 9", "Region 10")

alpha_diversity_region$US_region <- factor(alpha_diversity_region$US_region, levels = region_order)

# Boxplot for Shannon Index by region
ggplot(alpha_diversity_region, aes(x = US_region, y = shannon, group = US_region)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Shannon Diversity Index by Region",
       x = "US Region",
       y = "Shannon Index") +
  ylim(2,6)

# Boxplot for Simpson Index by region
ggplot(alpha_diversity_region, aes(x = US_region, y = simpson, group = US_region)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Simpson Diversity Index by Region",
       x = "US Region",
       y = "Simpson Index") +
  ylim(0.7, 1)


```
Measures of beta diversity v. time/region/cluster
Here looking at cg-mlst profiles and to determine diversity within Cluster 1 and 2 and cgmlst profile diversity
across time, region and cluster. Here  using Jaccard distance presence/absence of unique alleles from the 436 
unique alleles identified

```{r Clusters v. time/region/cluster, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}

library(vegan)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(tidyr)

#intermediate file for here... matricies complicating rownames... 
write.csv(cgmlst_clean, "allele_data.csv", row.names = TRUE)


pa_data <- data.frame(Isolate_ID = allele_data[[1]])
rm(allele_data)
#reimport allele_data with rownames as first column 

for (allele in names(allele_data)) {
  
  unique_variants <- unique(allele_data[[allele]][allele_data[[allele]] > 0])
  
  for (variant in unique_variants) {
    new_col_name <- paste(allele, "allele", variant, sep = "_")
    pa_data[[new_col_name]] <- ifelse(allele_data[[allele]] == variant, 1, 0)
  }
}

meta_data <-clean_combined_data

meta_data <- meta_data %>% select(ID, year, cluster_id, US_region)

meta_data <- meta_data %>% rename(Isolate_ID = ID)

filtered_pa_data <- pa_data %>% 
  filter(Isolate_ID %in% meta_data$Isolate_ID)

filtered_pa_data[is.na(filtered_pa_data)] <- 0

write.csv(filtered_pa_data, "allele_presence_absence.csv", row.names = FALSE)

#reimport allele_presence_absence data w/ rownames as first column

jaccard_dist <- vegdist(allele_presence_absence, method = "jaccard", na.rm = TRUE)

jaccard_dist

pcoa_jaccard <- cmdscale(jaccard_dist, eig = TRUE, k = 2)

pcoa_jaccard_df <- as.data.frame(pcoa_jaccard$points)
colnames(pcoa_jaccard_df) <- c("PC1", "PC2")

pcoa_jaccard_df$Isolate_ID <- rownames(pcoa_jaccard_df)

pcoa_metadata <- merge(pcoa_jaccard_df, meta_data, by = "Isolate_ID")

pcoa_metadata_filtered <- pcoa_metadata %>% filter(cluster_id != "Singleton")

my_colors <- c('Cluster 1'= '#63BB70', 'Cluster 2' = '#7CCEFD', 'Cluster 3' = '#FF5252', 'Cluster 4' = '#9574CD', 'Singleton' = '#58585A', 'Cluster 5' = '#FFF175', 'Cluster 6' = '#F44ED8', 'Cluster 7' = '#F77D55')

ggplot(pcoa_metadata_filtered, aes(x = PC1, y = PC2, color = cluster_id)) +
  geom_point() +
  stat_ellipse(type = "t", linetype = 2, level = 0.95)  +
  theme_minimal() +
  labs(titel = "PCoA of Jaccard Distance by Cluster",
  x = "PC1",
  y =  "PC2") + 
  scale_color_manual(values = my_colors)

#No differences between Region vs. Year....

region_colors <- c('Region 1'= '#63BB70', 'Region 2' = '#7CCEFD', 'Region 3' = '#FF5252', 'Region 4' = '#9574CD', 'Region 5' = '#58585A', 'Region 6' = '#FFF175', 'Region 7' = '#F44ED8', 'Region 8' = '#F77D55', 'Region 9' = '#F77D55', 'Region 10' = '#F77D55' )

ggplot(pcoa_metadata_filtered, aes(x = PC1, y = PC2, color = US_region)) +
  geom_point() +
  stat_ellipse(type = "t", linetype = 2, level = 0.95)  +
  theme_minimal() +
  labs(titel = "PCoA of Jaccard Distance by Region",
       x = "PC1",
       y =  "PC2", 
       color = "US Region") + 
  scale_color_manual(values = region_colors)


permanova_result <- adonis2(jaccard_dist ~ cluster_id, data = meta_data, method = "jaccard")

# View PERMANOVA results
permanova_result

write.csv(meta_data, "meta_data.csv", row.names = FALSE)



{r, leave one out appoarch for logistic regression analysis, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE }

library(tidyverse)
library(broom)
library(ggplot2)
library(brglm2)
library(car)
library(gt)
library(patchwork)

logit_campy_data <- clean_combined_data %>%
  mutate(
    Cluster_1 = ifelse(cluster_id == "Cluster 1", 1, 0),
    Cluster_2 = ifelse(cluster_id == "Cluster 2", 1, 0),
    Cluster_3 = ifelse(cluster_id == "Cluster 3", 1, 0),
    Cluster_4 = ifelse(cluster_id == "Cluster 4", 1, 0),
    Cluster_5 = ifelse(cluster_id == "Cluster 5", 1, 0),
    Cluster_6 = ifelse(cluster_id == "Cluster 6", 1, 0),
    Cluster_7 = ifelse(cluster_id == "Cluster 7", 1, 0),
    Singleton = ifelse(cluster_id == "Singleton", 1, 0),
    Region_1 = ifelse(US_region == "Region 1", 1, 0),
    Region_2 = ifelse(US_region == "Region 2", 1, 0),
    Region_3 = ifelse(US_region == "Region 3", 1, 0),
    Region_4 = ifelse(US_region == "Region 4", 1, 0),
    Region_5 = ifelse(US_region == "Region 5", 1, 0),
    Region_6 = ifelse(US_region == "Region 6", 1, 0),
    Region_7 = ifelse(US_region == "Region 7", 1, 0),
    Region_8 = ifelse(US_region == "Region 8", 1, 0),
    Region_9 = ifelse(US_region == "Region 9", 1, 0),
    Region_10 = ifelse(US_region == "Region 10", 1, 0)
  )

write.csv(logit_campy_data, "logit_data_check.csv", row.names = FALSE)

# data is formated correctly...

# building models...including brglm2 which is a package to that helps to reduce bias in logistic regression
# deals with complete or quas-complete seperation

Cluster_1_models <- list(
  Region_1 = glm(Cluster_1 ~ Region_1, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_2 = glm(Cluster_1 ~ Region_2, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_3 = glm(Cluster_1 ~ Region_3, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_4 = glm(Cluster_1 ~ Region_4, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_5 = glm(Cluster_1 ~ Region_5, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_6 = glm(Cluster_1 ~ Region_6, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_7 = glm(Cluster_1 ~ Region_7, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_8 = glm(Cluster_1 ~ Region_8, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_9 = glm(Cluster_1 ~ Region_9, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_10 = glm(Cluster_1 ~ Region_10, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit")
)


Cluster_2_models <- list(
  Region_1 = glm(Cluster_2 ~ Region_1, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_2 = glm(Cluster_2 ~ Region_2, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_3 = glm(Cluster_2 ~ Region_3, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_4 = glm(Cluster_2 ~ Region_4, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_5 = glm(Cluster_2 ~ Region_5, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_6 = glm(Cluster_2 ~ Region_6, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_7 = glm(Cluster_2 ~ Region_7, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_8 = glm(Cluster_2 ~ Region_8, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_9 = glm(Cluster_2 ~ Region_9, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_10 = glm(Cluster_2 ~ Region_10, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit")
)


Cluster_3_models <- list(
  Region_1 = glm(Cluster_3 ~ Region_1, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_2 = glm(Cluster_3 ~ Region_2, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_3 = glm(Cluster_3 ~ Region_3, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_4 = glm(Cluster_3 ~ Region_4, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_5 = glm(Cluster_3 ~ Region_5, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_6 = glm(Cluster_3 ~ Region_6, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_7 = glm(Cluster_3 ~ Region_7, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_8 = glm(Cluster_3 ~ Region_8, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_9 = glm(Cluster_3 ~ Region_9, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_10 = glm(Cluster_3 ~ Region_10, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit")
)


Cluster_4_models <- list(
  Region_1 = glm(Cluster_4 ~ Region_1, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_2 = glm(Cluster_4 ~ Region_2, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_3 = glm(Cluster_4 ~ Region_3, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_4 = glm(Cluster_4 ~ Region_4, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_5 = glm(Cluster_4 ~ Region_5, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_6 = glm(Cluster_4 ~ Region_6, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_7 = glm(Cluster_4 ~ Region_7, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_8 = glm(Cluster_4 ~ Region_8, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_9 = glm(Cluster_4 ~ Region_9, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_10 = glm(Cluster_4 ~ Region_10, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit")
)


Cluster_5_models <- list(
  Region_1 = glm(Cluster_5 ~ Region_1, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_2 = glm(Cluster_5 ~ Region_2, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_3 = glm(Cluster_5 ~ Region_3, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_4 = glm(Cluster_5 ~ Region_4, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_5 = glm(Cluster_5 ~ Region_5, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_6 = glm(Cluster_5 ~ Region_6, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_7 = glm(Cluster_5 ~ Region_7, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_8 = glm(Cluster_5 ~ Region_8, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_9 = glm(Cluster_5 ~ Region_9, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_10 = glm(Cluster_5 ~ Region_10, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit")
)


Cluster_6_models <- list(
  Region_1 = glm(Cluster_6 ~ Region_1, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_2 = glm(Cluster_6 ~ Region_2, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_3 = glm(Cluster_6 ~ Region_3, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_4 = glm(Cluster_6 ~ Region_4, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_5 = glm(Cluster_6 ~ Region_5, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_6 = glm(Cluster_6 ~ Region_6, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_7 = glm(Cluster_6 ~ Region_7, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_8 = glm(Cluster_6 ~ Region_8, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_9 = glm(Cluster_6 ~ Region_9, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_10 = glm(Cluster_6 ~ Region_10, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit")
)


Cluster_7_models <- list(
  Region_1 = glm(Cluster_7 ~ Region_1, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_2 = glm(Cluster_7 ~ Region_2, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_3 = glm(Cluster_7 ~ Region_3, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_4 = glm(Cluster_7 ~ Region_4, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_5 = glm(Cluster_7 ~ Region_5, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_6 = glm(Cluster_7 ~ Region_6, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_7 = glm(Cluster_7 ~ Region_7, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_8 = glm(Cluster_7 ~ Region_8, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_9 = glm(Cluster_7 ~ Region_9, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_10 = glm(Cluster_7 ~ Region_10, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit")
)

Singleton_models <- list(
  Region_1 = glm(Singleton ~ Region_1, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_2 = glm(Singleton ~ Region_2, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_3 = glm(Singleton~ Region_3, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_4 = glm(Singleton ~ Region_4, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_5 = glm(Singleton ~ Region_5, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_6 = glm(Singleton ~ Region_6, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_7 = glm(Singleton ~ Region_7, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_8 = glm(Singleton ~ Region_8, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_9 = glm(Singleton ~ Region_9, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit"),
  Region_10 = glm(Singleton ~ Region_10, family = binomial(link = "logit"), data = logit_campy_data, method = "brglmFit")
)



format_p_value <- function(x) {
  ifelse(x < 0.001, format(x, scientific = TRUE, digits = 2), scales::number(x, accuracy = 0.001))
}

cluster_1_summaries <- lapply(names(Cluster_1_models), function(region) {
  tidy(Cluster_1_models[[region]], conf.int = TRUE) %>%
    mutate(Cluster = "Cluster 1", Region = gsub("_", " ", region))
}) %>%
  bind_rows() %>%
  filter(term != "(Intercept)") %>%
  mutate(
    Odds_Ratio = exp(estimate),
    CI = paste0("(", round(exp(conf.low), 2), ", ", round(exp(conf.high), 2), ")"),
    p.value = format_p_value(p.value)
  )

cluster_2_summaries <- lapply(names(Cluster_2_models), function(region) {
  tidy(Cluster_2_models[[region]], conf.int = TRUE) %>%
    mutate(Cluster = "Cluster 2", Region = gsub("_", " ", region))
}) %>%
  bind_rows() %>%
  filter(term != "(Intercept)") %>%
  mutate(
    Odds_Ratio = exp(estimate),
    CI = paste0("(", round(exp(conf.low), 2), ", ", round(exp(conf.high), 2), ")"),
    p.value = format_p_value(p.value)
  )

cluster_3_summaries <- lapply(names(Cluster_3_models), function(region) {
  tidy(Cluster_3_models[[region]], conf.int = TRUE) %>%
    mutate(Cluster = "Cluster 3", Region = gsub("_", " ", region))
}) %>%
  bind_rows() %>%
  filter(term != "(Intercept)") %>%
  mutate(
    Odds_Ratio = exp(estimate),
    CI = paste0("(", round(exp(conf.low), 2), ", ", round(exp(conf.high), 2), ")"),
    p.value = format_p_value(p.value)
  )


cluster_4_summaries <- lapply(names(Cluster_4_models), function(region) {
  tidy(Cluster_4_models[[region]], conf.int = TRUE) %>%
    mutate(Cluster = "Cluster 4", Region = gsub("_", " ", region))
}) %>%
  bind_rows() %>%
  filter(term != "(Intercept)") %>%
  mutate(
    Odds_Ratio = exp(estimate),
    CI = paste0("(", round(exp(conf.low), 2), ", ", round(exp(conf.high), 2), ")"),
    p.value = format_p_value(p.value)
  )


cluster_5_summaries <- lapply(names(Cluster_5_models), function(region) {
  tidy(Cluster_5_models[[region]], conf.int = TRUE) %>%
    mutate(Cluster = "Cluster 5", Region = gsub("_", " ", region))
}) %>%
  bind_rows() %>%
  filter(term != "(Intercept)") %>%
  mutate(
    Odds_Ratio = exp(estimate),
    CI = paste0("(", round(exp(conf.low), 2), ", ", round(exp(conf.high), 2), ")"),
    p.value = format_p_value(p.value)
  )

cluster_6_summaries <- lapply(names(Cluster_5_models), function(region) {
  tidy(Cluster_6_models[[region]], conf.int = TRUE) %>%
    mutate(Cluster = "Cluster 6", Region = gsub("_", " ", region))
}) %>%
  bind_rows() %>%
  filter(term != "(Intercept)") %>%
  mutate(
    Odds_Ratio = exp(estimate),
    CI = paste0("(", round(exp(conf.low), 2), ", ", round(exp(conf.high), 2), ")"),
    p.value = format_p_value(p.value)
  )


cluster_7_summaries <- lapply(names(Cluster_5_models), function(region) {
  tidy(Cluster_7_models[[region]], conf.int = TRUE) %>%
    mutate(Cluster = "Cluster 7", Region = gsub("_", " ", region))
}) %>%
  bind_rows() %>%
  filter(term != "(Intercept)") %>%
  mutate(
    Odds_Ratio = exp(estimate),
    CI = paste0("(", round(exp(conf.low), 2), ", ", round(exp(conf.high), 2), ")"),
    p.value = format_p_value(p.value)
  )


singleton_summaries <- lapply(names(Singleton_models), function(region) {
  tidy(Singleton_models[[region]], conf.int = TRUE) %>%
    mutate(Cluster = "Cluster 7", Region = gsub("_", " ", region))
}) %>%
  bind_rows() %>%
  filter(term != "(Intercept)") %>%
  mutate(
    Odds_Ratio = exp(estimate),
    CI = paste0("(", round(exp(conf.low), 2), ", ", round(exp(conf.high), 2), ")"),
    p.value = format_p_value(p.value)
  )


Region_colors <- c("Region 1" = "#85755dff", 
                   "Region 2" = "#a8486cff", 
                   "Region 3" = "#393d6cff", 
                   "Region 4" = "#ca4848ff",
                   "Region 5" = "#317788ff", 
                   "Region 6"  = "#64780bff", 
                   "Region 7" = "#fcae17ff", 
                   "Region 8" = "#76518fff", 
                   "Region 9" =  "#932924ff", 
                   "Region 10" = "#127ab5ff"
                   )

cluster_1_summaries <- cluster_1_summaries %>%
  mutate(term = gsub("_", " ", term))

cluster_2_summaries <- cluster_2_summaries %>%
  mutate(term = gsub("_", " ", term))

cluster_3_summaries <- cluster_3_summaries %>%
  mutate(term = gsub("_", " ", term))

cluster_4_summaries <- cluster_4_summaries %>%
  mutate(term = gsub("_", " ", term))

cluster_5_summaries <- cluster_5_summaries %>%
  mutate(term = gsub("_", " ", term))

cluster_6_summaries <- cluster_6_summaries %>%
  mutate(term = gsub("_", " ", term))

cluster_7_summaries <- cluster_7_summaries %>%
  mutate(term = gsub("_", " ", term))

singleton_summaries <- singleton_summaries %>%
  mutate(term = gsub("_", " ", term))


create_forest_plot <- function(data, cluster_name) {

  data <- data %>%
    mutate(p.value = as.numeric(p.value))
  
  data %>%
    filter(p.value < 0.05) %>%
    ggplot(aes(x = reorder(term, estimate), y = exp(estimate), color = term)) +
    geom_point() +
    geom_hline(yintercept = 1, linetype = "dashed") +
    geom_errorbar(aes(ymin = exp(conf.low), ymax = exp(conf.high)), width = 0.2) +
    coord_flip() +
    labs(y = "Odds Ratio", title = paste("Forest Plot for", cluster_name)) +
    theme_minimal() + 
    theme(legend.position = "none", axis.title.y = element_blank(), axis.ticks.y = element_blank()) +
    scale_y_continuous(limits = c(0, 5.5)) +
    scale_color_manual(values = Region_colors)
}

cluster_1_plot <- create_forest_plot(cluster_1_summaries, "Cluster 1")
cluster_2_plot <- create_forest_plot(cluster_2_summaries, "Cluster 2")
cluster_3_plot <- create_forest_plot(cluster_3_summaries, "Cluster 3")
cluster_4_plot <- create_forest_plot(cluster_4_summaries, "Cluster 4")
cluster_5_plot <- create_forest_plot(cluster_5_summaries, "Cluster 5")
cluster_6_plot <- create_forest_plot(cluster_6_summaries, "Cluster 6")
cluster_7_plot <- create_forest_plot(cluster_7_summaries, "Cluster 7")
singleton_plot <- create_forest_plot(singleton_summaries, "Singleton")

Checking Plots
print(cluster_1_plot)  # Cluster 1-3 are the only ones with more than 1 significant association

combined_plot <- (cluster_1_plot | cluster_2_plot)


print(combined_plot)


cluster_total <- logit_campy_data %>%
  summarise(
    Cluster_1 = sum(Cluster_1),
    Cluster_2 = sum(Cluster_2),
    Cluster_3 = sum(Cluster_3),
    Cluster_4 = sum(Cluster_4),
    Cluster_5 = sum(Cluster_5),
    Cluster_6 = sum(Cluster_6),
    Cluster_7 = sum(Cluster_7),
    Singleton = sum(Singleton)
  )

region_totals_cluster_1 <- logit_campy_data %>%
  filter(Cluster_1 == 1) %>%
  summarise(across(starts_with("Region"), sum)) %>%
  pivot_longer(cols = everything(), names_to = "Region", values_to = "Total") %>%
  mutate(Region = gsub("_", " ", Region)) 
  
  

cluster_1_summaries <- cluster_1_summaries %>%
  mutate(term = gsub("_", " ", term)) %>%
  left_join(region_totals_cluster_1, by = c("term" = "Region")) %>%
  mutate(term = paste0(term, " (n=", Total, ")")) %>%
  select(-Total)


region_totals_cluster_2 <- logit_campy_data %>%
  filter(Cluster_2 == 1) %>%
  summarise(across(starts_with("Region"), sum)) %>%
  pivot_longer(cols = everything(), names_to = "Region", values_to = "Total") %>%
  mutate(Region = gsub("_", " ", Region)) 



cluster_2_summaries <- cluster_2_summaries %>%
  mutate(term = gsub("_", " ", term)) %>%
  left_join(region_totals_cluster_2, by = c("term" = "Region")) %>%
  mutate(term = paste0(term, " (n=", Total, ")")) %>%
  select(-Total)




##Change gt for cluster 2

gt_cluster_table <- cluster_1_summaries %>%
  select(term, Odds_Ratio, CI, p.value) %>%
  gt() %>%
  tab_header(
    title = md(paste("Cluster 1 Analysis (n=", cluster_total$Cluster_1, ")", sep = ""))
  ) %>%
  cols_label(
    term = "Region",
    Odds_Ratio = "Odds Ratio",
    CI = "95% CI",
    p.value = "P-Value"
  ) %>%
  tab_spanner(
    label = "Cluster Analysis",
    columns = c(Odds_Ratio, CI, p.value)
  ) %>%
  tab_options(
    table.font.names = "sans-serif",
    table.font.size = px(12)
  )


gt_cluster_table_2 <- cluster_2_summaries %>%
  select(term, Odds_Ratio, CI, p.value) %>%
  gt() %>%
  tab_header(
    title = md(paste("Cluster 2 Analysis (n=", cluster_total$Cluster_2, ")", sep = ""))
  ) %>%
  cols_label(
    term = "Region",
    Odds_Ratio = "Odds Ratio",
    CI = "95% CI",
    p.value = "P-Value"
  ) %>%
  tab_spanner(
    label = "Cluster Analysis",
    columns = c(Odds_Ratio, CI, p.value)
  ) %>%
  tab_options(
    table.font.names = "sans-serif",
    table.font.size = px(12)
  )


print(gt_cluster_table_2)


{r, viruelence factors by sourec, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE }

library(tidyverse)
library(dplyr)
library(ggplot2)

# of virulence genes violin plot 

v_data <- virulence_results %>%
  rename(AccessionNum = 1)%>%
  mutate(AccessionNum = str_extract(AccessionNum, "TW\\d+"))

human_tw_n <- human_datatable$Isolates
cattle_tw_n <- cattle_datatable$Isolates

v_data <- v_data %>%
  mutate(Source = case_when(
    AccessionNum %in% human_tw_n ~ "Human",
    AccessionNum %in% cattle_tw_n ~ "Cattle",
    TRUE ~ "Unknown"))

v_counts <- v_data %>%
  group_by(AccessionNum, Source) %>%
  summarise(Virulence_gene_count = n(), .groups = 'drop')

write.csv(v_counts, "v_counts.csv", row.names = FALSE)

ggplot(v_counts, aes(x = Source, y = Virulence_gene_count, fill = Source)) +
  geom_violin(trim = FALSE) + 
  geom_boxplot(width = 0.2, outlier.shape = NA) +
  geom_point(position = position_jitter(width = 0.2), shape = 21, size = 3, color = "black", fill = "white") +
  scale_fill_manual(values = c("Cattle" = "#79bd8f", "Human" = "#70dfec")) +
  labs(title = "Number of Virulence Genes per Source",
       x = "Source",
       y = "Number of Virulence Genes") +
  theme_minimal()

t_test_result <- t.test(Virulence_gene_count ~ Source, data = v_counts)
print(t_test_result)

wilcox_test_result <- wilcox.test(Virulence_gene_count ~ Source, data = v_counts)
print(wilcox_test_result)


unique_product_names <- unique(v_data$PRODUCT)

print(unique_product_names)

#creating functional classes for virulence gene data... 

v_data <- v_data %>%
  mutate(Function_class = case_when(
    str_detect(PRODUCT, "Flagella") ~ "Flagella",
    str_detect(PRODUCT, "Capsule") ~ "Capsule",
    str_detect(PRODUCT, "LOS") ~ "LOS",
    str_detect(PRODUCT, "CDT") ~ "CDT",
    str_detect(PRODUCT, "chemotaxis") ~ "Chemotaxis",
    str_detect(PRODUCT, "Type IV secretion system") ~ "Type IV Secretion System",
    str_detect(PRODUCT, "adhesin|fimbriae|pilus") ~ "Adhesins",
    TRUE ~ "Other"
  ))

write.csv(v_data, "check_virulence.csv", row.names = FALSE)

class_v_counts <- v_data %>%
  group_by(Source, Function_class) %>%
  summarise(Count = n(), .groups = 'drop')

total_counts <- class_v_counts %>%
  group_by(Function_class) %>%
  summarise(Total_count = sum(Count), .groups = 'drop')

class_v_percent <- class_v_counts %>%
  left_join(total_counts, by = "Function_class") %>%
  mutate(Percent = (Count/Total_count) *100) %>%
  select(-Total_count)

ggplot(class_v_percent, aes(x = Function_class, y = Percent, fill = Source)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("Cattle" = "#79bd8f", "Human" = "#70dfec")) +
  labs(title = "Distribution of Functional Classes of Virulence Genes by Source",
       x = "Functional Class",
       y = "Percentage of Virulence Genes") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

contingency_tables <- class_v_percent %>%
  pivot_wider(names_from = Source, values_from = c(Count, Percent), values_fill = list(Count = 0, Percent = 0)) %>%
  mutate(Total = Count_Human + Count_Cattle)

total_counts_source <- v_data %>%
  group_by(Source) %>%
  summarise(Total_Count = n(), .groups = 'drop')

total_human <- total_counts_source %>% filter(Source == "Human") %>% pull(Total_Count)
total_cattle <- total_counts_source %>% filter(Source == "Cattle") %>% pull(Total_Count)


chi_square_results <- contingency_tables %>%
  rowwise() %>%
  mutate(
    Chi_Square_Test = list(
      chisq.test(matrix(c(Count_Human, total_human - Count_Human, Count_Cattle, total_cattle - Count_Cattle), nrow = 2))
    )
  ) %>%
  mutate(
    Chi_Square_Statistic = Chi_Square_Test$statistic,
    Chi_Square_P_Value = Chi_Square_Test$p.value,
    Chi_Square_Expected = list(Chi_Square_Test$expected)
  ) %>%
  select(Function_class, Chi_Square_Statistic, Chi_Square_P_Value, Chi_Square_Expected)

chi_square_summary <- chi_square_results %>%
  mutate(Significant = Chi_Square_P_Value < 0.05) %>%
  select(Function_class, Chi_Square_Statistic, Chi_Square_P_Value, Significant)

print(chi_square_summary)


#scatter plot of the genes... 

fill_color <- "#c3aed6"

gene_counts <- v_data %>%
  group_by(GENE, Source) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = Source, values_from = Count, values_fill = list(Count = 0)) %>%
  rename(Cattle_count = Cattle, Human_count = Human)


ggplot(gene_counts, aes(x = Human_count, y = Cattle_count)) +
  geom_point(color = "black", fill = fill_color, size = 5, shape = 21) +
  labs(title = "Virulence Gene Counts by Source",
       x = "Count in Human Genomes",
       y =  "Count in Cattle Genomes") + 
  theme_minimal() +
  scale_x_continuous(limits = c(0,250)) +
  scale_y_continuous(limits = c(0, 110))


write.csv(gene_counts, "virulence_gene_counts.csv", row.names = FALSE)


{r, plasmid data, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE }

library(tidyverse)
library(dplyr)
library(ggplot2)

#  of plasmids per genome 

plasmids <- merged_mobtyper_results 

plasmids <- plasmids %>%
  mutate(TW_number = str_extract(sample_id, "TW\\d+")) %>%
  group_by(TW_number) %>%
  summarise(Plasmid_count = n()) %>%
  ungroup()

human_isolates <- human_datatable %>%
  select(Isolates) %>%
  mutate(Source = "Human")

cattle_isolates <- cattle_datatable %>%
  select(Isolates) %>%
  mutate(Source = "Cattle")

combined_isolates <- bind_rows(human_isolates, cattle_isolates) %>%
  rename(TW_number = Isolates)

plasmids_data <- combined_isolates %>%
  left_join(plasmids, by = "TW_number") %>%
  replace_na(list(Plasmid_count = 0))

my_colors <- c(Cattle = "#79bd8f", Human = "#70dfec")

ggplot(plasmids_data, aes(x = Source, y = Plasmid_count, fill = Source)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.2, outlier.shape = NA) +
  geom_point(position = position_jitter(width = 0.2), shape = 21, size = 3, color = "black", fill = "white") +
  scale_fill_manual(values = my_colors) +
  labs(title = "Number of Plasmids in Human vs. Cattle Isolates",
       x = "Source",
       y = "Number of Plasmids") +
  theme_minimal() +
  ylim(0, 5)

t_test_result <- t.test(Plasmid_count ~ Source, data = plasmids_data)
print(t_test_result)

wilcox_test_result <- wilcox.test(Plasmid_count ~ Source, data = plasmids_data)
print(wilcox_test_result)

plasmids_tree_data <- merged_mobtyper_results

plasmids_tree_data <- plasmids_tree_data %>%
  mutate(TW_number = str_extract(sample_id, "TW\\d+"))


plasmids_tree_data <- combined_isolates %>%
  left_join(plasmids_tree_data, by = "TW_number")

plasmids_tree_data <- plasmids_tree_data %>%
  mutate(
    pVIR = ifelse(primary_cluster_id %in% c("AE190", "AE015"), 1, 0)
  ) %>%
  mutate(
    p118 = ifelse(primary_cluster_id == "AC439", 1, 0 )
  ) %>%
  mutate(
    pCJDM = ifelse(primary_cluster_id == "AC321", 1, 0)
  ) %>%
  mutate(
    pTET = ifelse(primary_cluster_id== "AC320", 1, 0)
  )

out_data <- plasmids_tree_data %>%
  select(TW_number, Source, predicted_mobility, primary_cluster_id, pVIR, p118, pCJDM, pTET)

write.csv(out_data, "plasmids_presence_absence.csv", row.names = FALSE)

Figure in ITOL for the rest...

